Download and Save the Actual Data on MCI

### 2024-01-21 outputs/paper/paper.qmd
figure explanation corrected

```{r}
#| message: false
#| echo: false
# Load packages
library(opendatatoronto)
library(janitor)
library(tidyverse)
library(leaflet)
library(dplyr)
library(here)
library(knitr)
```

@@

## Data
### Data Source and Software
The data utilized throughout this paper, in csv format, is obtained from Toronto Police Services on the City of Toronto Open Data Website [@rOpenDataToronto], with the title “MAJOR CRIME INDICATORS”.

@@

### Data Analysis and Visualizations

```{r}
#| message: false
#| echo: false
#| label: tbl-clean_mci_data
#| tbl-cap: The Number of Major Crimes and the Month with the Most Crimes (2014-2023)
# Create a table with year, 2014 to 2023, the number of crimes in each year, and the month with the most crimes correspondingly.
clean_mci_data |>
  filter(occ_year >= 2014) |>
  group_by(occ_year, occ_month) |>
  summarise(n = n()) |>
  group_by(occ_year) |>
  summarise(
    total_number_of_crimes = sum(n),
    the_most_frequent_month = occ_month[which.max(n)]
  ) |>
  distinct(occ_year, .keep_all = TRUE) |>
  select(occ_year, total_number_of_crimes, the_most_frequent_month) |> 
  kable(
    col.names = c("Year", "Total Number of Crimes", "Month with the Most Crimes"),
    booktabs = TRUE
  )
```

```{r}
#| message: false
#| echo: false
#| label: fig-types_of_crimes
#| fig-cap: The Number of Crimes Reported, Classified by Crime Types (2014-2023)
clean_mci_data |>
  ggplot(mapping = aes(x = factor(report_year), fill = mci_category)) +
  geom_bar(position = "stack", stat = "count") +
  labs(x = "Year", y = "Total Number of Crimes", fill = "Types of Major Crimes") +
  theme(legend.position = "right") +
  scale_fill_brewer(palette = "Set3")
```

@@

```{r}
#| include: false
clean_mci_data <- clean_mci_data |> 
  mutate(
    weekday_weekend =
      case_match(
        occ_dow,
        "Wednesday" ~ "Weekday",
        "Tuesday" ~ "Weekday",
        "Thursday" ~ "Weekday",
        "Sunday" ~ "Weekend",
        "Monday" ~ "Weekday",
        "Friday" ~ "Weekday",
        "Saturday" ~ "Weekend",
        "" ~ "Other"))
head(clean_mci_data)
```

```{r}
#| message: false
#| echo: false
#| label: fig-area_of_crimes
#| fig-cap: Total Number of Crimes in Different Police Division Areas
clean_mci_data |>
  filter(occ_year >= 2014) |>
  count(occ_year, division_area) |>
  ggplot(aes(x = occ_year, y = n, color = division_area)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Year",
       y = "Total Number of Crimes",
       color = "Division Area")
```

```{r}
#| message: false
#| echo: false
#| label: fig-time_of_crimes
#| fig-cap: Number of Crimes in a Day
clean_mci_data |>
  filter(weekday_weekend == 'Weekday' | weekday_weekend == 'Weekend') |>
  count(occ_hour, weekday_weekend) |>
  ggplot(mapping = aes(x = occ_hour, y = n, color = weekday_weekend)) +
  geom_line() +
  theme(legend.position = "top") +
  labs(x = "Time In a Day", y = "Number of Crimes", color = "Weekday or Weekend") +
  scale_color_manual(values = c("Weekday" = "steelblue3", "Weekend" = "indianred3"))
```

>>>>

```{r}
#| message: false
#| echo: false
# Load packages
library(opendatatoronto)
library(janitor)
library(tidyverse)
library(leaflet)
library(dplyr)
library(ggplot2)
library(here)
library(knitr)
```

@@

## Data

### Data Source and Software
The data utilized throughout this paper, in csv format, is obtained from Toronto Police Services on the City of Toronto Open Data Website [@rOpenDataToronto], with the title “MAJOR CRIME INDICATORS”.

@@

### Data Analysis and Visualizations
The Major Crimes in Toronto dataset contains 27 columns and 372899 rows, including the repeated occurrences with different crime categories. All the samples of major crimes were reported from 2014 to 2023. Among 27 columns, the variables I will use are x_id, report_year, occ_year, occ_month, occ_dow (Day of the Week Offence Occurred), occ_hour, mci_category, division. I will create two more variables, division_area, which classifies 18 police divisions (including a NSA) into 6 regions of Toronto (including Other for NSA), and weekend_weekday, which indicates whether the occurrence happen on a weekday or a weekend.

```{r}
#| message: false
#| echo: false
#| label: tbl-clean_mci_data
#| tbl-cap: The Number of Major Crimes and the Month with the Most Crimes (2014-2023)
# Create a table with year, 2014 to 2023, the number of crimes in each year, and the month with the most crimes correspondingly.
clean_mci_data |>
  filter(occ_year >= 2014) |>
  group_by(occ_year, occ_month) |>
  summarise(n = n()) |>
  group_by(occ_year) |>
  summarise(
    total_number_of_crimes = sum(n),
    the_most_frequent_month = occ_month[which.max(n)]
  ) |>
  distinct(occ_year, .keep_all = TRUE) |>
  select(occ_year, total_number_of_crimes, the_most_frequent_month) |> 
  kable(
    col.names = c("Year", "Total Number of Crimes", "Month with the Most Crimes"),
    booktabs = TRUE
  )
```
Table 1 provides a brief summary of the total number of crimes from 2014 to 2023. Over this period, the overall number of crimes shows a consistent increase, except for the years 2020 and 2021, which have a decline probably due to the impact of the pandemic. However, in 2022 and 2023, there is a significant increase by more than 6000. In each year, the month with the highest number of crimes tends to be either July, August, or October, concentrating primarily in the summer-fall period.

```{r}
#| message: false
#| echo: false
#| label: fig-types_of_crimes
#| fig-cap: The Number of Crimes Reported, Classified by Crime Types (2014-2023)
clean_mci_data |>
  ggplot(mapping = aes(x = factor(report_year), fill = mci_category)) +
  geom_bar(position = "stack", stat = "count") +
  labs(x = "Year", y = "Total Number of Crimes", fill = "Types of Major Crimes") +
  theme(legend.position = "right") +
  scale_fill_brewer(palette = "Set3")
```
Figure 1 visualizes the annual distribution of major crime types with five categories: Assault, Break and Enter, Auto Theft, Robbery, and Theft Over. Across all years, 'Assault' records the highest count, followed by 'Break and Enter' and 'Auto Theft'. 'Robbery' and 'Theft Over' show comparatively lower numbers. Notably, there has been a slight decrease in “Break and Enter” since the pandemic, and an obvious increase in 'Auto Theft' during 2022 and 2023, leading ‘Auto Theft’ surpassing the number of 'Break and Enter’.

@@

```{r}
#| include: false
clean_mci_data <- clean_mci_data |> 
  mutate(
    weekday_weekend =
      case_match(
        occ_dow,
        "Wednesday" ~ "Weekday",
        "Tuesday" ~ "Weekday",
        "Thursday" ~ "Weekday",
        "Sunday" ~ "Weekend",
        "Monday" ~ "Weekday",
        "Friday" ~ "Weekday",
        "Saturday" ~ "Weekend",
        "" ~ "Other"))
clean_mci_data
```

```{r}
#| message: false
#| echo: false
#| label: fig-area_of_crimes
#| fig-cap: Total Number of Crimes in Different Police Division Areas
clean_mci_data |>
  filter(occ_year >= 2014) |>
  count(occ_year, division_area) |>
  ggplot(aes(x = occ_year, y = n, color = division_area)) +
  geom_point() +
  geom_jitter() +
  geom_smooth(method = lm, se = FALSE) +
  labs(x = "Year",
       y = "Total Number of Crimes",
       color = "Division Area")
```
There are 18 police divisions, which are D11, D12, D13, D14, D22, D23, D31, D32, D33, D41, D42, D43, D51, D52, D53, D54, D55, and NSA. Putting them into 6 regions based on geographical location, Etobicoke-York, Etobicoke, North York, Scarborough, Toronto-East York, and other for NSA.
Based on Figure 2, Toronto-East York reports the highest amount of crimes, and Etobicoke-York has the lowest. North York, Scarborough, and Etobicoke-York have similar numbers, but the rate of increase in North York is higher than that observed in the other two.

Based on Figure 2, Toronto-East York reports the highest amount of crimes, and Etobicoke-York has the lowest. North York, Scarborough, and Etobicoke-York have similar numbers, but the rate of increase in North York is higher than that observed in the other two.

```{r}
#| message: false
#| echo: false
#| label: fig-time_of_crimes
#| fig-cap: Number of Crimes in a Day
clean_mci_data |>
  filter(weekday_weekend == 'Weekday' | weekday_weekend == 'Weekend') |>
  count(occ_hour, weekday_weekend) |>
  ggplot(mapping = aes(x = occ_hour, y = n, color = weekday_weekend)) +
  geom_line() +
  theme(legend.position = "top") +
  labs(x = "Time In a Day", y = "Number of Crimes", color = "Weekday or Weekend") +
  scale_color_manual(values = c("Weekday" = "steelblue3", "Weekend" = "indianred3"))
```
Figure 3 indicates the preferred time for committing crimes. The graph reveals a higher likelihood of criminal behaviors during the evening, with occurrences steadily increasing as the night progresses towards midnight and afterwards, the occurrences decrease until six in the morning. Interestingly, there is a surprising local peak at noon, while 11:00 and 13:00 do not show a high probability.

### 2024-01-21 scripts/01-download_data.qmd
figure explanation corrected

```{r}
# Install all the packages I need for this analysis
install.packages('opendatatoronto')
install.packages('dplyr')
install.packages('tidyverse')
install.packages('janitor')
install.packages('leaflet')
install.packages('knitr')
install.packages('here')
```

>>>>

```{r}
# Install all the packages I need for this analysis
install.packages('opendatatoronto')
install.packages('dplyr')
install.packages('tidyverse')
install.packages('janitor')
install.packages('leaflet')
install.packages('ggplot2')
install.packages('knitr')
install.packages('here')
```
